Untitled

### 2021-02-01 ms/manuscript.Rmd
Update MS

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology.
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located \circa{ }15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (Figs. `r figure_num("habit")`, `r s_figure_num("tree")`).
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006 for Hymenophyllaceae.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (Figs. `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

@@

# FIXME: double check on these p-values
# ttest_pval("Polyphlebium_endlicherianum", "recovery")
# ttest_pval("Crepidomanes_minutum2", "recovery")
# ttest_pval("Hymenophyllum_digitatum", "recovery")
# ttest_pval("Hymenophyllum_polyanthos", "recovery")
# ttest_pval("Polyphlebium_borbonicum", "recovery")
```

Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases.
A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`a).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A. dentatum* after confirming its lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H.* *pallidum* and *P.* *borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

@@

@@

Relative humidity in the growth chambers was high when samples were first inserted, then decreased to the target value maintained by each drying salt (Figs. `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).

DT of gametophytes did not significantly exceed that of sporophytes for any species except for saxicolous *P. endlicherianum* (Fig. 5; *t*-test, *P* = ), and was significantly lower in gametophytes than sporophytes of some high elevation epiphytic species including *C. minutum* var. 2, *H.* *digitatum*, *H. polyanthos*, and *P. borbonicum* (`r figure("combined")`; *t*-test, *P* = XXX, XXX, XXX, and XXX, respectively).

The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`) while taking phylogeny into account. Sporophytes had `r sporo_dt_effect_mean` greater DT (95% CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).

@@

Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases (Appendix S4).
Unlike DT, photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes are adapted for higher light levels, whereas gametophytes are adapted for lower light levels (Fig. 5).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).
The GLMMs showed a significant effect of generation, but neither range type nor growth habit, on PPFD~95%~ and ETR~max~ (Table 3).

@@

Desiccation tolerance tended to increase with VPD for sporophytes, but not gametophytes (`r figure("corr")`).
These trends were marginally non-significant when analyzed using PGLS for sporophytes (*P* = `r sporo_vpd_pval`) and not significant for gametophytes (*P* = `r gameto_vpd_pval`; `r table("pgls")`).

No significant relationship was detected between range size of gametophytes and DT (`r figure("corr")`, `r table("pgls")`).

>>>>

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology.
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located \circa{ }15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`).
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006 for Hymenophyllaceae.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22 °C to 25 °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

@@

t_test_results

# Extract list of species with significant p-values for gameto DT less than sporo
dt_g_less_than_s <-
t_test_results %>% 
  filter(est_diff_sporo_gameto > 0, response == "recovery", p_value < 0.05) %>%
  abbrev_sp() %>%
  mutate(species = str_replace_all(species, "_", ". ")) %>%
  pull(species) %>%
  knitr::combine_words(before = "*")

# Extract p-values for species with gameto DT less than sporo
dt_g_less_than_s_pvals <-
t_test_results %>%
  filter(est_diff_sporo_gameto > 0, response == "recovery", p_value < 0.05) %>%
  pull(species) %>%
  map_chr(~ttest_pval(species_select = ., response_select = "recovery", data = t_test_results)) %>%
  knitr::combine_words()
```

Ability to recover from desiccation was tested in sporophytes of `r n_dt_species_sporo` species and gametophytes of `r n_dt_species_gameto` species, including `r n_dt_species_both` species with both life phases.

A general correlation of DT with habitat was observed in sporophytes (`r figure("dt")`a).
Both terrestrial species (*Abrodictyum dentatum* (Bosch) Ebihara & K Iwats. and *C. apiifolia*) lacked even moderate DT, and failed to recover from the gentlest treatment of 2 d at --38 NaCl (more stringent tests were not conducted on *A. dentatum* after confirming its lack of DT at this level).
Saxicolous species (*Crepidomanes kurzii* (Bedd.) Tagawa & K. Iwats., *Polyphlebium endlicherianum* (C. Presl) Ebihara & K. Iwats., and *Vandenboschia maxima* (Blume) Copel.) also had low DT; only *C. kurzii* was able to recover from desiccation up to --86 MPa for 2 d but not more.
All epiphytic species were capable of recovering from desiccation to a certain degree, but ability varied between species.
Most low elevation epiphytes could not tolerate long periods (15 d) of desiccation; only *Crepidomanes bipunctatum* (Poir.) Copel. was able to recover from desiccation longer than 2 d.
Most of the high elevation epiphytes could recover from 2 d treatments regardless of desiccation intensity (except for *H. pallidum* and *P. borbonicum*, which failed to recover from 2 d at --282 MPa), and many could withstand desiccation for 15 d.
Interestingly, *C. minutum* var. 2, *Hymenophyllum multifidum* (G. Forst.) Sw., and *Hymenophyllum polyanthos* (Sw.) Sw. could recover from desiccation at --86 or --282 MPa for 15 d, but not --38 MPa.

@@

DT of gametophytes did not significantly exceed that of sporophytes for any species, and was significantly lower in `r dt_g_less_than_s` (`r figure("combined")`; *t*-test, *P* = `r dt_g_less_than_s_pvals`, respectively).

@@

Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).

The GLMM including the effect of generation only was selected as the best-fitting model for DT (`r table("glmm")`). 
Sporophytes had `r sporo_dt_effect_mean` greater DT (95% CI `r sporo_dt_effect_lower` to `r sporo_dt_effect_upper`) relative to gametophytes (`r table("glmm-pars")`).

@@

Light responses were measured in sporophytes of `r n_light_species_sporo` species and gametophytes of `r n_light_species_gameto` species, including `r n_light_species_both` species with both life phases.

Photosynthetic optima tended to vary between life phases, but not across growth habits or range type: sporophytes were adapted for higher light levels, whereas gametophytes were adapted for lower light levels (`r figure("combined")`).
Sporophytes reached their maximum photosynthetic rate (PPFD~95%~) at light levels ca.
`r mean_par_sporo` µmol·m^--2^ ·s^--1^ , whereas gametophytes did so at `r mean_par_gameto` µmol·m^--2^ ·s^--1^.
Similarly, sporophytes attained higher rates of electron transport (ETR~max~; `r mean_etr_sporo` µmol·m^--2^ ·s^--1^) than gametophytes (`r mean_etr_gameto` µmol·m^--2^ ·s^--1^).

The GLMM including the effect of generation only was selected as the best-fitting model for ETR~max~, and growth habit + generation was selected for PPFD~95%~ (`r table("glmm")`).
However, in the GLMM for PPFD~95%~, only the generation parameter was significant (`r table("glmm-pars")`).

@@

Desiccation tolerance tended to increase with VPD for sporophytes, but not gametophytes (`r figure("corr")`).
This trend was marginally non-significant when analyzed using PGLS (*P* = `r sporo_vpd_pval`; `r table("pgls")`).

No significant relationship was detected between range size of gametophytes beyond sporophytes and DT (`r figure("corr")`, `r table("pgls")`).
